Technique for \earlywarning\ detection of compact binary coalescence and its implication for multi-messenger astronomy

### 2011-06-03 method.tex
more words
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@855 ccdd05d5-56c1-461a-bf98-85f6876523b4

The first transformation is to chop the time domain templates up into time
slices.  Since each template piece is disjoint in time, the resulting set for a
single template is orthogonal.  Given the chirp like structure of the
templates, the early time slices have significantly lower bandwidth and can be
safely downsampled.  The downsampling removes a factor of $\sim 10$ samples
overall and also allows the filters to be evaluated at about a $\sim 10$ times
lower rate.  This amounts to more than a factor of 100 in the floating point
operations per second required.  However, the resulting filters are still not
orthogonal across the mass parameter space, and are in fact highly redundant.
We use the singular value decomposition to produce an orthogonal filter set
from the time sliced templates~\cite{Cannon:2010p10398}.  We find that this
reduces the number of sample points that need to be filtered by another factor
of $\sim 100$.  The combined methods reduce the number of floating point
operations to the level where they are competitive with the conventional
matched filter approach.  In the remainder of this section we describe the
\lloid\ algorithm in detail and provide some basic computational cost scaling.

@@

Our first innovation and the first step of our proposed method
is to divide the templates into \emph{time slices} in a time-domain
analogue to the frequency-domain decomposition described in ~\cite{Marion2004,
Buskulic2010, beauville2006, beauville2008}\editorial{Four citations is way too many!  We should pick just \emph{one} citation for \textsc{mbta}, and it should be the most recent and complete description of it.}.  A matched filter is constructed
for each time slice.  The outputs form an ensemble of partial \SNR{}
streams.  By linearity, these partial \SNR\ streams can be suitably time
delayed and summed to reproduce the \SNR\ of the full template.  To wit,
%
\begin{equation}
\label{eq:time-slices}
h_{i}[k] = \sum_{s=0}^{S-1}
	\cases{%
		h_i^s[k] & for $t^s \leqslant k / f^0 < t^{s+1}$ \\
		0 & otherwise
	}
\end{equation}
%
for $S$ integers $\{f^0 t^s\}$ such that $0  = f^0 t^0 < f^0 t^1 < \cdots < f^0 t^S = N$.
We will show
in the next section that this, combined with the singular value docomposition,
is sufficient to enable a computationally efficient time-domain search and
furthermore is an essential part of an \earlywarning\ detection scheme.

>>>>

The first transformation is to chop the time domain templates up into time
slices.  Since each template piece is disjoint in time, the resulting set for a
single template is orthogonal.  Given the chirp like structure of the
templates, the early time slices have significantly lower bandwidth and can be
safely downsampled.  The downsampling removes a factor of $\sim 10$ samples on
average and also allows the filters to be evaluated at about a $\sim 10$ times
lower rate on average.  This amounts to more than a factor of 100 in the
floating point operations per second required.  However, the resulting filters
are still not orthogonal across the mass parameter space, and are in fact
highly redundant.  We use the singular value decomposition to produce an
orthogonal filter set from the time sliced templates~\cite{Cannon:2010p10398}.
We find that this reduces the number of sample points that need to be filtered
by another factor of $\sim 100$.  The combined methods reduce the number of
floating point operations to the level where they are competitive with the
conventional matched filter approach.  In the remainder of this section we
describe the \lloid\ algorithm in detail and provide some basic computational
cost scaling.

@@

The first step of our proposed method is to divide the templates into
\emph{time slices} in a time-domain analogue to the frequency-domain
decomposition described in ~\cite{Marion2004, Buskulic2010, beauville2006,
beauville2008}\editorial{Four citations is way too many!  We should pick just
\emph{one} citation for \textsc{mbta}, and it should be the most recent and
complete description of it.}.  A matched filter is constructed for each time
slice.  The outputs form an ensemble of partial \SNR{} streams.  By linearity,
these partial \SNR\ streams can be suitably time delayed and summed to
reproduce the \SNR\ of the full template.  To wit,
%
\begin{equation}
\label{eq:time-slices}
h_{i}[k] = \sum_{s=0}^{S-1}
	\cases{%
		h_i^s[k] & for $t^s \leqslant k / f^0 < t^{s+1}$ \\
		0 & otherwise
	}
\end{equation}
%
for $S$ integers $\{f^0 t^s\}$ such that $0  = f^0 t^0 < f^0 t^1 < \cdots < f^0 t^S = N$.
We will show
in the next section that this, combined with the singular value docomposition,
is sufficient to enable a computationally efficient time-domain search and
furthermore is an essential part of an \earlywarning\ detection scheme.
